\chapter{Procedures} 
\label{procedures}  
\section{Evacuating Transfer lines}

\subsection{Cryofab TLs}

\subsection{LTL}

\section{Indium Extrusion}

\section{Making Indium Seals}

\section{Breaking Indium Seals}
\begin{enumerate}
 \item Remove screws with 1.5 mm allen key and the modified MC wrench.  Visually inspect each screw for indium or signs of stripping, and turn each nut back on the screw it came off of.  If there is any torque resistance or sign of stripping, the screw was likely damaged when the indium seal was tightened and both the screw and nut should be discarded.  If there is no damage, return the nuts/screws to the bag labeled ``MC Screws and Nuts''.
\item Use two MC screws and the 1.5 mm allen key as jacking screws the break the indium seal.  Defining the top of the fridge (in the horizontal orientation) as 12:00, the jacking screw holes are at 1:00 and 5:00 look upstream.  The jacking screw holes can be identified by threaded holes in the MC bell housing that have no matching through holes on the fridge MC flange.  Alternate between screws making 1/4 turns until the MC is free.
\item Remove the MC from the fridge, careful not to damage the sensors on the copper dam.
\item Remove the jacking screws from the MC and place in the bag labeled ``MC Screws and Nuts''.
\item Scrape indium, if any, off the MC and place the MC in the safety zone.
\item Scrape indium off the fridge flange, recovering as much as possible in the indium scrap container.  \text{Be very careful not to touch the stainless steel beam window}.
\end{enumerate}


\section{Magnet Lead Installation}

\section{Mounting Fridge}

\section{Dismounting Fridge}

\section{Cleaning Viton/Buna O-rings}

\section{Changing Kenol Connectors}

\section{Installing Heating Tapes}

\section{Filling \lnn{} Trap}

\section{Purging Helium Lines}

\section{Vacuum Rise Testing}

\section{Stripping Fridge}

\section{Assembling Fridge}

  \subsection{MC}
  \subsection{Microwave Guide Support}
  \subsection{IVC}
  \subsection{OVC}

\section{Removing/Replacing \het{} Baffles}